---
title: "242 Final Project"
author: "Lizzy Kinnard and Shelley Facente"
date: "October 31, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#open packages
library(here)
library(foreign)
library(labelled)
library(dplyr)
```

```{r dataload, include=FALSE}
#set working directory
#setwd("~/Desktop/lizzyandshelley/Dataset") #LIZZY
setwd("~/GitHub/lizzyandshelley/Dataset")  #SHELLEY

#load the data
df <- read.spss("Ricky.SAV", use.value.labels = TRUE, to.data.frame=TRUE)

#get variable labels
attr(df, "variable.labels")
df.labels <- as.data.frame(attr(df, "variable.labels"))
QDSlabels <- data.frame(rownames(df.labels))
df.labels <- cbind(QDSlabels, df.labels)
names(df.labels) <- c("VAR_ID", "description")
rm(QDSlabels)

#select people with no LTFU
dfc <- subset(df, FU2==1 & FU1=="Yes")

#select variables
dfs <- dfc %>%
  select("GENDER", "AGE", "RACELG", "HOMELESS", "HOMELESS1", "HOMELESS2", "MOSHOME", "MOSHOME1", "MOSHOM2", "NVHOUSED", 
         "ALWYSHL", "HOUSED", "HLP6M", "HLP6M1", "HLP6M2", "NHLP6M", "NHLP6M1", "NHLP6M2", "NDLSRCF", "NDLSRCF1", 
         "NDLSRCF2", "PUBINJ", "PUBINJ1", "PUBINJ2", "IJWOTH", "IJWOTH1", "IJWOTH2", "IJNON", "IJNON1", "IJNON2", 
         "ODP6M", "ODP6M1", "ODP6M2", "ODTIMES1", "ODTIMES2", "WOD6M", "WOD6M1", "WOD6M2", "ODTRAIN1", "NALOWN1", 
         "NALAVAIL1", "POLICE1", "POLICE2", "POLTIME1", "ARREST1", "ARREST2", "JAIL1", "JAIL2", "TX6MOS1", "TX6MOS2")
```

```{r dataplay, include=FALSE}
#USE THIS CODE TO LOOK UP A VARIABLE DESCRIPTION
#df.labels$description[df.labels$VAR_ID=="GENDER"]

#inspect the data
names(df)
dim(df)
length(unique(dfs$SUBJECT))

table(dfs$GENDER, useNA = "ifany")
table(dfs$FU1, useNA = "ifany")
table(dfs$FU2, useNA = "ifany")


#look at homeless variables
table(dfs$HOMELESS, useNA = "ifany")
table(dfs$HOMELESS1, useNA = "ifany")
table(dfs$HOMELESS2, useNA = "ifany")

#look at homeless variables ricky created
table(dfs$NVHOUSED, useNA = "ifany")
table(dfs$ALWYSHL, useNA = "ifany")
table(dfs$HOUSED, useNA = "ifany")

#look at initiation variables
apply(dfs[c("HLP6M", "HLP6M1", "HLP6M2")], 2, table)

#look at arrest
table(dfs$ARREST2, useNA = "ifany")

```

Questions for Ricky:

1. C14 in the baseline survey is "Have you ever INJECTED someone for their first hit?" What variable name is that? We can only find those that apply to past 6 months (not ever).

2. Do you have a question for helping to inject someone ever, or is it only for the first hit in all cases? 

3. Why are the values for the "helping others to initiate" timepoints so odd? (SEE LINE 71)

4. You have a question in the 12 month survey for "ever arrested" and we assume that's why there so many NAs for the question about arrest in the past 6 months at the 12-month follow-up (ARREST2). a) this wasn't asked before the 12-month f/u, right? b) what is the variable name for the "ever arrested" question at 12-months? 

```{r analysis, include=TRUE}

```






