---
title: "ComparingBaselineLTFU"
author: "Elizabeth Kinnard"
date: "3/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval=T, warning = FALSE, message = FALSE, global.par=TRUE)
knitr::opts_knit$set(root.dir="~/Documents/GitHub/lizzyandshelley/Dataset")
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=85),tidy=TRUE)
```

Open packages.
```{r}
library(base)
library(knitr)
library(dplyr)
library(here)
library(foreign)
library(labelled)
library(dplyr)
library(reshape2)
library(tableone)
library(forcats)
library(geepack)
library(doBy)
library(blm)
library(lmtest)
library(foreign)
library(MASS)
```

Open working directory and load the dataset with all waves of CTC. 
```{r}
#set working directory
setwd("~/Documents/GitHub/lizzyandshelley/Dataset") #LIZZY

#load the data
df_allwaves <- read.spss("CTCAllWaves.SAV", use.value.labels = TRUE, to.data.frame=TRUE)
```

Get variable labels.
```{r, echo=F, eval=F}
#get variable labels
attr(df_allwaves, "variable.labels")
df_allwaves.labels <- as.data.frame(attr(df_allwaves, "variable.labels"))
QDSlabels <- data.frame(rownames(df_allwaves.labels))
df_allwaves.labels <- cbind(QDSlabels, df_allwaves.labels)
names(df_allwaves.labels) <- c("VAR_ID", "description")
rm(QDSlabels)

#USE THIS CODE TO LOOK UP A VARIABLE DESCRIPTION
#df_allwaves.labels$description[df_allwaves.labels$VAR_ID=="GENDER"]
```

See who showed up to FU1 and FU2.
```{r}
table(df_allwaves$FU1, useNA = "ifany")
table(df_allwaves$FU2, useNA = "ifany")
```

Create variables to compare those who would eventually show up to 12-month visit (FU2) to those who would not (the comparison itself is at baseline visit).
```{r}
#city
table(df_allwaves$LOCALID, useNA = "ifany")

#gender (combine other and transgender into other so 3 cat variable)
df_allwaves$GENDER <- df_allwaves$GENDER %>% fct_collapse("Other" = c("Transgender","Other"))
table(df_allwaves$GENDER, useNA = "ifany")

#age
table(df_allwaves$AGE, useNA = "ifany")
table(df_allwaves$agecat, useNA = "ifany") #(<30, 30 to 39, 40-49, 50 or more) 

#race (recoded into fewer categories)
table(df_allwaves$RACELG, useNA = "ifany")
df_allwaves$RACE_cat <- df_allwaves$RACELG %>% fct_collapse("Other" = c("Native American","Asian/Pacific Islander", "Mixed Race/Other"))

df_allwaves$RACE_cat <- droplevels(df_allwaves$RACE_cat, exclude=c("Don't know", "Refused"))
table(df_allwaves$RACE_cat, useNA = "ifany")

#education (High school education or more)
table(df_allwaves$ED, useNA = "ifany")

#homelessness (Consider Self Homeless)
table(df_allwaves$HOMELESS, useNA = "ifany")

#incarceration (Any jail in the last 6 months)
#not possible to use because they didn't ask about jail at baseline


###SPEEDBALL###
table(df_allwaves$SB30, useNA = "ifany") #used speedball yes/no (heroin + cocaine or crack)
table(df_allwaves$SBNINJ, useNA = "ifany")  #any non-injection speedball
table(df_allwaves$SBX, useNA = "ifany")  #non-injection speedball (# of times)
table(df_allwaves$SBINJ, useNA = "ifany")  #any injection speedball 
table(df_allwaves$SBIX, useNA = "ifany")  #injection speedball (# of times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$SBX <- as.integer(df_allwaves$SBX) #change to numeric instead of factor
    df_allwaves$SBIX <- as.integer(df_allwaves$SBIX) #change to numeric instead of factor
    
    df_allwaves$SB.freq <-   ifelse(df_allwaves$SBX>=30 | df_allwaves$SBIX>=30, 2, ifelse((df_allwaves$SBX>0 & df_allwaves$SBX<30) & (df_allwaves$SBIX>0 & df_allwaves$SBIX<30), 1, 
                                                                   ifelse(df_allwaves$SBX==0 & df_allwaves$SBIX==0, 0, NA)))
    df_allwaves$SB.freq[is.na(df_allwaves$SB.freq)] <- 0
    table(df_allwaves$SB.freq, useNA = "ifany")
    
###GOOFBALL###
table(df_allwaves$GB30, useNA = "ifany")   #used goofball yes/no (heroin + meth)
table(df_allwaves$GBNINJ, useNA = "ifany")   #any non-injection goofball 
table(df_allwaves$GBX, useNA = "ifany")   #non-injection goofball (# times)
table(df_allwaves$GBINJ, useNA = "ifany")   #any injection goofball
table(df_allwaves$GBIX, useNA = "ifany")   #injection goofball (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$GBX <- as.integer(df_allwaves$GBX) #change to numeric instead of factor
    df_allwaves$GBIX <- as.integer(df_allwaves$GBIX) #change to numeric instead of factor
    
    df_allwaves$GB.freq <-   ifelse(df_allwaves$GBX>=30 | df_allwaves$GBIX>=30, 2, ifelse((df_allwaves$GBX>0 & df_allwaves$GBX<30) & (df_allwaves$GBIX>0 & df_allwaves$GBIX<30), 1, 
                                                                   ifelse(df_allwaves$GBX==0 & df_allwaves$GBIX==0, 0, NA)))
    df_allwaves$GB.freq[is.na(df_allwaves$GB.freq)] <- 0
    table(df_allwaves$GB.freq, useNA = "ifany")

###CRACK###
table(df_allwaves$CRCK30, useNA = "ifany")  #used crack yes/no
table(df_allwaves$CRKNINJ, useNA = "ifany") #any non-injection crack
table(df_allwaves$CRCKX, useNA = "ifany")   #non-injection crack (# times)
table(df_allwaves$CRKINJ, useNA = "ifany") #any injection crack 
table(df_allwaves$CRCKIX, useNA = "ifany") #injection crack (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$CRCKX <- as.integer(df_allwaves$CRCKX) #change to numeric instead of factor
    df_allwaves$CRCKIX <- as.integer(df_allwaves$CRCKIX) #change to numeric instead of factor
    
    df_allwaves$CRCK.freq <-   ifelse(df_allwaves$CRCKX>=30 | df_allwaves$CRCKIX>=30, 2, ifelse((df_allwaves$CRCKX>0 & df_allwaves$CRCKX<30) & (df_allwaves$CRCKIX>0 & df_allwaves$CRCKIX<30), 1, 
                                                                   ifelse(df_allwaves$CRCKX==0 & df_allwaves$CRCKIX==0, 0, NA)))
    df_allwaves$CRCK.freq[is.na(df_allwaves$CRCK.freq)] <- 0
    table(df_allwaves$CRCK.freq, useNA = "ifany")

###COCAINE###
table(df_allwaves$PCOKE30, useNA = "ifany") #used powder cocaine yes/no
table(df_allwaves$CCNINJ, useNA = "ifany") #any non-injection powder cocaine 
table(df_allwaves$PCOKEX, useNA = "ifany") #non-injection powder cocaine (times)
table(df_allwaves$CCINJ, useNA = "ifany")  #any injection powder cocaine
table(df_allwaves$PCOKEIX, useNA = "ifany") #injection powder cocaine (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$PCOKEX <- as.integer(df_allwaves$PCOKEX) #change to numeric instead of factor
    df_allwaves$PCOKEIX <- as.integer(df_allwaves$PCOKEIX) #change to numeric instead of factor
    
    df_allwaves$PCOKE.freq <-   ifelse(df_allwaves$PCOKEX>=30 | df_allwaves$PCOKEIX>=30, 2, ifelse((df_allwaves$PCOKEX>0 & df_allwaves$PCOKEX<30) & (df_allwaves$PCOKEIX>0 & df_allwaves$PCOKEIX<30), 1, 
                                                                   ifelse(df_allwaves$PCOKEX==0 & df_allwaves$PCOKEIX==0, 0, NA)))
    df_allwaves$PCOKE.freq[is.na(df_allwaves$PCOKE.freq)] <- 0
    table(df_allwaves$PCOKE.freq, useNA = "ifany")
    
###METHAMPHETAMINE###
table(df_allwaves$METH30, useNA = "ifany") #used meth yes/no
table(df_allwaves$MTNINJ, useNA = "ifany") #any non-injection meth 
table(df_allwaves$METHX, useNA = "ifany") #non-injection meth (# times)
table(df_allwaves$MTINJ, useNA = "ifany") #any injection meth
table(df_allwaves$METHIX, useNA = "ifany") #injection meth (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$METHX <- as.integer(df_allwaves$METHX) #change to numeric instead of factor
    df_allwaves$METHIX <- as.integer(df_allwaves$METHIX) #change to numeric instead of factor
    
    df_allwaves$METH.freq <-   ifelse(df_allwaves$METHX>=30 | df_allwaves$METHIX>=30, 2, ifelse((df_allwaves$METHX>0 & df_allwaves$METHX<30) & (df_allwaves$METHIX>0 & df_allwaves$METHIX<30), 1, 
                                                                   ifelse(df_allwaves$METHX==0 & df_allwaves$METHIX==0, 0, NA)))
    df_allwaves$METH.freq[is.na(df_allwaves$METH.freq)] <- 0
    table(df_allwaves$METH.freq, useNA = "ifany")

###NONMEDICAL PRESCRIPTION STIMULANT###
table(df_allwaves$ST30, useNA = "ifany") #used NMPS yes/no
table(df_allwaves$STNINJ, useNA = "ifany") #any non-injection NMPS
table(df_allwaves$STX, useNA = "ifany") #non-injection NMPS (# times)
table(df_allwaves$STINJ, useNA = "ifany") #any injection NMPS
table(df_allwaves$STIX, useNA = "ifany") #injection NMPS (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$STX <- as.integer(df_allwaves$STX) #change to numeric instead of factor
    df_allwaves$STIX <- as.integer(df_allwaves$STIX) #change to numeric instead of factor
    
    df_allwaves$ST.freq <-   ifelse(df_allwaves$STX>=30 | df_allwaves$STIX>=30, 2, ifelse((df_allwaves$STX>0 & df_allwaves$STX<30) & (df_allwaves$STIX>0 & df_allwaves$STIX<30), 1, 
                                                                   ifelse(df_allwaves$STX==0 & df_allwaves$STIX==0, 0, NA)))
    df_allwaves$ST.freq[is.na(df_allwaves$ST.freq)] <- 0
    table(df_allwaves$ST.freq, useNA = "ifany")

###HEROIN###
table(df_allwaves$HER30, useNA = "ifany") #used heroin yes/no
table(df_allwaves$HRNINJ, useNA = "ifany") #any non-injection heroin 
table(df_allwaves$HERX, useNA = "ifany") #non-injection heroin (# times)
table(df_allwaves$HRINJ, useNA = "ifany") #any injection heroin 
table(df_allwaves$HERIX, useNA = "ifany") #injection heroin (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$HERX <- as.integer(df_allwaves$HERX) #change to numeric instead of factor
    df_allwaves$HERIX <- as.integer(df_allwaves$HERIX) #change to numeric instead of factor
    
    df_allwaves$HER.freq <-   ifelse(df_allwaves$HERX>=30 | df_allwaves$HERIX>=30, 2, ifelse((df_allwaves$HERX>0 & df_allwaves$HERX<30) & (df_allwaves$HERIX>0 & df_allwaves$HERIX<30), 1, 
                                                                  ifelse(df_allwaves$HERX==0 & df_allwaves$HERIX==0, 0, NA)))
    df_allwaves$HER.freq[is.na(df_allwaves$HER.freq)] <- 0
    table(df_allwaves$HER.freq, useNA = "ifany")

###NONMEDICAL PRESCRIPTION OPIOIDS###
table(df_allwaves$OP30, useNA = "ifany") #used NMPO yes/no
table(df_allwaves$OPNINJ, useNA = "ifany") #any non-injection NMPO 
table(df_allwaves$OPX, useNA = "ifany") #non-injection NMPO (# times)
table(df_allwaves$OPINJ, useNA = "ifany") #any injection NMPO
table(df_allwaves$OPIX, useNA = "ifany") #injection NMPO (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$OPX <- as.integer(df_allwaves$OPX) #change to numeric instead of factor
    df_allwaves$OPIX <- as.integer(df_allwaves$OPIX) #change to numeric instead of factor
    
    df_allwaves$OP.freq <-   ifelse(df_allwaves$OPX>=30 | df_allwaves$OPIX>=30, 2, ifelse((df_allwaves$OPX>0 & df_allwaves$OPX<30) & (df_allwaves$OPIX>0 & df_allwaves$OPIX<30), 1, 
                                                                   ifelse(df_allwaves$OPX==0 & df_allwaves$OPIX==0, 0, NA)))
    df_allwaves$OP.freq[is.na(df_allwaves$OP.freq)] <- 0
    table(df_allwaves$OP.freq, useNA = "ifany")

###METHADONE###
table(df_allwaves$MD30, useNA = "ifany") #used methadone yes/no
table(df_allwaves$MDNINJ, useNA = "ifany") #any non-injection methadone
table(df_allwaves$MDX, useNA = "ifany") #non-injection methadone (# times)
table(df_allwaves$MDINJ, useNA = "ifany") #any injection methadone 
table(df_allwaves$MDIX, useNA = "ifany") #injection methadone (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$MDX <- as.integer(df_allwaves$MDX) #change to numeric instead of factor
    df_allwaves$MDIX <- as.integer(df_allwaves$MDIX) #change to numeric instead of factor
    
    df_allwaves$MD.freq <-   ifelse(df_allwaves$MDX>=30 | df_allwaves$MDIX>=30, 2, ifelse((df_allwaves$MDX>0 & df_allwaves$MDX<30) & (df_allwaves$MDIX>0 & df_allwaves$MDIX<30), 1, 
                                                                   ifelse(df_allwaves$MDX==0 & df_allwaves$MDIX==0, 0, NA)))
    df_allwaves$MD.freq[is.na(df_allwaves$MD.freq)] <- 0
    table(df_allwaves$MD.freq, useNA = "ifany")
    
###BUPRENORPHINE###
table(df_allwaves$BUP30, useNA = "ifany") #used bup yes/no
table(df_allwaves$BUPNINJ, useNA = "ifany") #any non-injection bup 
table(df_allwaves$BUPX, useNA = "ifany") #non-injection bup (# times)
table(df_allwaves$BUPINJ, useNA = "ifany") #any injection bup 
table(df_allwaves$BUPIX, useNA = "ifany") #injection bup (# times)

    #creating frequency variable (2= greater than or equal to 30 times, 1 = greater than 0 but less than 30, 0 = none)
    df_allwaves$BUPX <- as.integer(df_allwaves$BUPX) #change to numeric instead of factor
    df_allwaves$BUPIX <- as.integer(df_allwaves$BUPIX) #change to numeric instead of factor
    
    df_allwaves$BUP.freq <-   ifelse(df_allwaves$BUPX>=30 | df_allwaves$BUPIX>=30, 2, ifelse((df_allwaves$BUPX>0 & df_allwaves$BUPX<30) & (df_allwaves$BUPIX>0 & df_allwaves$BUPIX<30), 1, 
                                                                   ifelse(df_allwaves$BUPX==0 & df_allwaves$BUPIX==0, 0, NA)))
    df_allwaves$BUP.freq[is.na(df_allwaves$BUP.freq)] <- 0
    table(df_allwaves$BUP.freq, useNA = "ifany")

#FREQUENCY OF STIMULANT USE
#includes crack cocaine, powder cocaine, meth, nonmedical rx stimulants, speedball, and goofball
df_allwaves$STIM.freq <- ifelse(df_allwaves$SB.freq==2 | df_allwaves$GB.freq==2 | df_allwaves$CRCK.freq==2 | df_allwaves$PCOKE.freq==2 | df_allwaves$METH.freq==2 | df_allwaves$ST.freq==2, 2, ifelse(df_allwaves$SB.freq==0 & df_allwaves$GB.freq==0 & df_allwaves$CRCK.freq==0 & df_allwaves$PCOKE.freq==0 & df_allwaves$METH.freq==0 & df_allwaves$ST.freq==0, 0, 1))

df_allwaves$STIM.freq <- as.factor(df_allwaves$STIM.freq)
table(df_allwaves$STIM.freq, useNA = "ifany")


#FREQUENCY OF OPIOID USE
#includes speedball, goofball, heroin, nonmedical rx opioids, methadone, bup
df_allwaves$OPIOID.freq <- ifelse(df_allwaves$SB.freq==2 | df_allwaves$GB.freq==2 | df_allwaves$HER.freq==2 | df_allwaves$OP.freq==2 | df_allwaves$MD.freq==2 | df_allwaves$BUP.freq==2, 2, ifelse(df_allwaves$SB.freq==0 & df_allwaves$GB.freq==0 & df_allwaves$HER.freq==0 & df_allwaves$OP.freq==0 & df_allwaves$MD.freq==0 & df_allwaves$BUP.freq==0, 0, 1))

df_allwaves$OPIOID.freq <- as.factor(df_allwaves$OPIOID.freq)
table(df_allwaves$OPIOID.freq, useNA = "ifany")

####ANY STIMULANT USE IN PAST 30 DAYS (injection and/or non-injection composite)###
#includes crack cocaine, powder cocaine, meth, nonmedical rx stimulants, speedball, and goofball
df_allwaves$ANYSTIMUSE <- ifelse(df_allwaves$CRKNINJ=="Yes" | df_allwaves$CCNINJ=="Yes" | df_allwaves$MTNINJ=="Yes" | df_allwaves$STNINJ=="Yes" |df_allwaves$SBNINJ =="Yes" |df_allwaves$GBNINJ=="Yes" | df_allwaves$CRKINJ=="Yes" | df_allwaves$CCINJ=="Yes" | df_allwaves$MTINJ=="Yes" | df_allwaves$STINJ=="Yes" |df_allwaves$SBINJ=="Yes"| df_allwaves$GBINJ=="Yes", 1, ifelse(df_allwaves$CRKNINJ=="No" &df_allwaves$CCNINJ=="No" & df_allwaves$MTNINJ=="No" & df_allwaves$STNINJ=="No" & df_allwaves$SBNINJ=="No" & df_allwaves$GBNINJ=="No" & df_allwaves$CRKINJ=="No" &  df_allwaves$CCINJ=="No" & df_allwaves$MTINJ=="No" & df_allwaves$STINJ=="No" & df_allwaves$SBINJ=="No" & df_allwaves$GBINJ=="No", 0, NA))

df_allwaves$ANYSTIMUSE <- as.factor(df_allwaves$ANYSTIMUSE)
table(df_allwaves$ANYSTIMUSE, useNA = "ifany")

###ANY OPIOID USE IN PAST 30 DAYS (injection and/or non-injection composite)###
#includes speedball, goofball, heroin, nonmedical rx opioids, methadone, bup
df_allwaves$ANYOPIOIDUSE <- ifelse(df_allwaves$SBINJ=="Yes" | df_allwaves$GBINJ=="Yes" | df_allwaves$MDINJ=="Yes" | df_allwaves$BUPINJ=="Yes" | df_allwaves$HRINJ=="Yes" | df_allwaves$OPINJ=="Yes" |df_allwaves$SBNINJ=="Yes" | df_allwaves$GBNINJ=="Yes" |  df_allwaves$MDNINJ=="Yes" |  df_allwaves$BUPNINJ=="Yes" |  df_allwaves$HRNINJ=="Yes" | df_allwaves$OPNINJ=="Yes", 1, ifelse(df_allwaves$SBINJ=="No" & df_allwaves$GBINJ=="No"  & df_allwaves$MDINJ=="No"  & df_allwaves$BUPINJ=="No" & df_allwaves$HRINJ=="No" & df_allwaves$OPINJ=="No"  & df_allwaves$SBNINJ=="No" & df_allwaves$GBNINJ=="No"  & df_allwaves$MDNINJ=="No"  & df_allwaves$BUPNINJ=="No" & df_allwaves$HRNINJ=="No" & df_allwaves$OPNINJ=="No", 0, NA))

df_allwaves$ANYOPIOIDUSE <- as.factor(df_allwaves$ANYOPIOIDUSE)
table(df_allwaves$ANYOPIOIDUSE, useNA = "ifany")

#Any OAT in the past six months (IF METHADONE DETOX = YES OR METHADONE MAINTENANCE = YES OR BUP = YES)
df_allwaves$OAT6month <- ifelse(df_allwaves$DETOX=="Yes" | df_allwaves$MAINTN=="Yes" | df_allwaves$BUPETX=="Yes", 1, 0)
table(df_allwaves$OAT6month, useNA = "ifany")

#overdose witnessed (witnessed an overdose in the last 6 months)
table(df_allwaves$WOD6M, useNA = "ifany")

#personal overdose history (Overdosed in the last 6 months)
table(df_allwaves$ODP6M, useNA = "ifany")

#####syringe access variables

#Most Frequent Needle Source Last 6 Mos.
table(df_allwaves$NDLSRCF, useNA = "ifany") #this has too many categories

#RECODING Most Frequent Needle Source Last 6 Mos (combine all other categories except needle exchange program into one)
df_allwaves$needleexchange <- df_allwaves$NDLSRCF %>% fct_collapse("Other" = c("Needle distribution program or outreach worker","From someone who goes to the needle exchange", "Bought from a pharmacy", "Bought or got from an unauthorized sources (street, a friend)", "My own prescription for syringes", "Shooting gallery or drug dealer", "Other"))

#Most Frequent Needle Source Last 6 Mos (RECODED Needle exchange program vs other)
#USE THIS ONE IN THE TABLE BELOW 
table(df_allwaves$needleexchange, useNA = "ifany")

#Number of new needles obtained for your own use in the last 30 days
table(df_allwaves$NMNDLES, useNA = "ifany")
class(df_allwaves$NMNDLES)

#need to do this so I don't lose the zero values of NMNDLES
##USE THIS ONE IN THE TABLE BELOW (will compare means)
df_allwaves$NMNDLES <- as.character.factor(df_allwaves$NMNDLES)
df_allwaves$NMNDLES <- as.numeric(df_allwaves$NMNDLES)

#if we end up needing to create a categorical variable to compare categories
#df_allwaves$NMNDLES.freq <- ifelse(df_allwaves$NMNDLES==0, 0, ifelse(df_allwaves$NMNDLES>0 & df_allwaves$NMNDLES<100, 1, ifelse(df_allwaves$NMNDLES>=100 & df_allwaves$NMNDLES <200, 2, ifelse(df_allwaves$NMNDLES>=200 & df_allwaves$NMNDLES <300, 3, ifelse(df_allwaves$NMNDLES>=300 & df_allwaves$NMNDLES<400, 4, ifelse(df_allwaves$NMNDLES>=400 & df_allwaves$NMNDLES<500, 5, ifelse(df_allwaves$NMNDLES>=500 & df_allwaves$NMNDLES<600, 6, 7)))))))

#checking that the categories turned out correctly    
#table(df_allwaves$NMNDLES.freq, useNA = "ifany")
```

Create a table to compare those who would eventually show up to 12-month visit (FU2 = 1) to those who would not (FU2 = 0).
```{r}
# Identify the variables of interest
myVars <- c("LOCALID", "GENDER","agecat","RACE_cat", "ED","HOMELESS", "STIM.freq", "OPIOID.freq", "OAT6month", "WOD6M", "ODP6M", "needleexchange", "NMNDLES") 

# Specify which of the variables are categorical
catVars <- c("LOCALID", "GENDER", "agecat", "RACE_cat", "ED","HOMELESS", "STIM.freq", "OPIOID.freq","OAT6month", "WOD6M", "ODP6M", "needleexchange") 

# Create an overall table 1
#REMINDER THAT TABLE1 PACKAGE DOES WEIRD STUFF WITH MISSING, SO CHECK N (PERCENTAGE) WITH CROSSTABS 

table1.overall <- CreateTableOne(vars= myVars, 
                                 data= df_allwaves, 
                                 factorVars = catVars)

#this is if we want table 1 not stratified by any outcome (eg FU2)
#print(table1.overall, showAllLevels = TRUE, missing = FALSE)

table1.byFU2 <- CreateTableOne(vars=myVars, 
                                   data=df_allwaves, 
                                   factorVars = catVars,
                                   strata="FU2")

print(table1.byFU2, showAllLevels = TRUE, test=TRUE, missing = FALSE)

#check column percentages of bivariates. this is because prop.table will show missing percentages whereas tableone package doesn't include the missing data when calculating column percents.
#table(df_allwaves$X, df_allwaves$FU2, useNA = "ifany")
#round(prop.table(table(df_allwaves$X, df_allwaves$FU2, useNA = "ifany"),2),3)
```





