---
title: "Opioid and stimulant use variable construction"
author: "Elizabeth Kinnard"
date: "10/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Open packages.
```{r}
library(base)
library(knitr)
library(dplyr)
library(here)
library(foreign)
library(labelled)
library(dplyr)
library(reshape2)
library(tableone)
library(forcats)
library(geepack)
library(doBy)
library(blm)
library(lmtest)
library(foreign)
library(MASS)
```

Open working directory and load/label the data.
```{r}
#set working directory
setwd("~/Documents/GitHub/lizzyandshelley/Dataset") #LIZZY

#load the data
df <- read.spss("CTC12month.SAV", use.value.labels = TRUE, to.data.frame=TRUE)
```

Get variable labels.
```{r, echo=F, eval=F}
#get variable labels
attr(df, "variable.labels")
df.labels <- as.data.frame(attr(df, "variable.labels"))
QDSlabels <- data.frame(rownames(df.labels))
df.labels <- cbind(QDSlabels, df.labels)
names(df.labels) <- c("VAR_ID", "description")
rm(QDSlabels)

#USE THIS CODE TO LOOK UP A VARIABLE DESCRIPTION
#df.labels$description[df.labels$VAR_ID=="GENDER"]
```

Subset the data to delete all rows that are entirely NA (since they didn't show up to 12 month visit).
```{r}
df <- df[rowSums(is.na(df)) < 600,] #rows reduced from N=984 to N=536 people

#EXAMPLES OF SUBSETTING
#df <- subset(df, SUMINJ>0)
#df <- subset(df, IJWOTH_3cat>0) 
```

Create frequency variables for all the different stimulants and opioids.
```{r}
###SPEEDBALL###
table(df$SB30, useNA = "ifany") #used speedball yes/no (heroin + cocaine or crack)
table(df$SBNINJ, useNA = "ifany")  #any non-injection speedball
table(df$SBX, useNA = "ifany")  #non-injection speedball (# of times)
table(df$SBINJ, useNA = "ifany")  #any injection speedball 
table(df$SBIX, useNA = "ifany")  #injection speedball (# of times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$SBX <- as.integer(df$SBX) #change to numeric instead of factor
    df$SBIX <- as.integer(df$SBIX) #change to numeric instead of factor
    
    df$SB_dailymore <- ifelse(df$SBX>=30 | df$SBIX>=30, 1, 0)
    table(df$SB_dailymore, useNA = "ifany")


###GOOFBALL###
table(df$GB30, useNA = "ifany")   #used goofball yes/no (heroin + meth)
table(df$GBNINJ, useNA = "ifany")   #any non-injection goofball 
table(df$GBX, useNA = "ifany")   #non-injection goofball (# times)
table(df$GBINJ, useNA = "ifany")   #any injection goofball
table(df$GBIX, useNA = "ifany")   #injection goofball (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$GBX <- as.integer(df$GBX) #change to numeric instead of factor
    df$GBIX <- as.integer(df$GBIX) #change to numeric instead of factor
    
    df$GB_dailymore <- ifelse(df$GBX>=30 | df$GBIX>=30, 1, 0)
    table(df$GB_dailymore, useNA = "ifany")

    
###CRACK###
table(df$CRCK30, useNA = "ifany")  #used crack yes/no
table(df$CRKNINJ, useNA = "ifany") #any non-injection crack
table(df$CRCKX, useNA = "ifany")   #non-injection crack (# times)
table(df$CRKINJ, useNA = "ifany") #any injection crack 
table(df$CRCKIX, useNA = "ifany") #injection crack (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$CRCKX <- as.integer(df$CRCKX) #change to numeric instead of factor
    df$CRCKIX <- as.integer(df$CRCKIX) #change to numeric instead of factor
    
    df$CRCK_dailymore <- ifelse(df$CRCKX>=30 | df$CRCKIX>=30, 1, 0)
    table(df$CRCK_dailymore, useNA = "ifany")

 
###COCAINE###
table(df$PCOKE30, useNA = "ifany") #used powder cocaine yes/no
table(df$CCNINJ, useNA = "ifany") #any non-injection powder cocaine 
table(df$PCOKEX, useNA = "ifany") #non-injection powder cocaine (times)
table(df$CCINJ, useNA = "ifany")  #any injection powder cocaine
table(df$PCOKEIX, useNA = "ifany") #injection powder cocaine (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$PCOKEX <- as.integer(df$PCOKEX) #change to numeric instead of factor
    df$PCOKEIX <- as.integer(df$PCOKEIX) #change to numeric instead of factor
    
    df$PCOKE_dailymore <- ifelse(df$PCOKEX>=30 | df$PCOKEIX>=30, 1, 0)
    table(df$PCOKE_dailymore, useNA = "ifany")

    
###METHAMPHETAMINE###
table(df$METH30, useNA = "ifany") #used meth yes/no
table(df$MTNINJ, useNA = "ifany") #any non-injection meth 
table(df$METHX, useNA = "ifany") #non-injection meth (# times)
table(df$MTINJ, useNA = "ifany") #any injection meth
table(df$METHIX, useNA = "ifany") #injection meth (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$METHX <- as.integer(df$METHX) #change to numeric instead of factor
    df$METHIX <- as.integer(df$METHIX) #change to numeric instead of factor
    
    df$METH_dailymore <- ifelse(df$METHX>=30 | df$METHIX>=30, 1, 0)
    table(df$METH_dailymore, useNA = "ifany")


###NONMEDICAL PRESCRIPTION STIMULANT###
table(df$ST30, useNA = "ifany") #used NMPS yes/no
table(df$STNINJ, useNA = "ifany") #any non-injection NMPS
table(df$STX, useNA = "ifany") #non-injection NMPS (# times)
table(df$STINJ, useNA = "ifany") #any injection NMPS
table(df$STIX, useNA = "ifany") #injection NMPS (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$STX <- as.integer(df$STX) #change to numeric instead of factor
    df$STIX <- as.integer(df$STIX) #change to numeric instead of factor
    
    df$ST_dailymore <- ifelse(df$STX>=30 | df$STIX>=30, 1, 0)
    table(df$ST_dailymore, useNA = "ifany")


###HEROIN###
table(df$HER30, useNA = "ifany") #used heroin yes/no
table(df$HRNINJ, useNA = "ifany") #any non-injection heroin 
table(df$HERX, useNA = "ifany") #non-injection heroin (# times)
table(df$HRINJ, useNA = "ifany") #any injection heroin 
table(df$HERIX, useNA = "ifany") #injection heroin (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$HERX <- as.integer(df$HERX) #change to numeric instead of factor
    df$HERIX <- as.integer(df$HERIX) #change to numeric instead of factor
    
    df$HER_dailymore <- ifelse(df$HERX>=30 | df$HERIX>=30, 1, 0)
    table(df$HER_dailymore, useNA = "ifany")


###NONMEDICAL PRESCRIPTION OPIOIDS###
table(df$OP30, useNA = "ifany") #used NMPO yes/no
table(df$OPNINJ, useNA = "ifany") #any non-injection NMPO 
table(df$OPX, useNA = "ifany") #non-injection NMPO (# times)
table(df$OPINJ, useNA = "ifany") #any injection NMPO
table(df$OPIX, useNA = "ifany") #injection NMPO (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$OPX <- as.integer(df$OPX) #change to numeric instead of factor
    df$OPIX <- as.integer(df$OPIX) #change to numeric instead of factor
    
    df$OP_dailymore <- ifelse(df$OPX>=30 | df$OPIX>=30, 1, 0)
    table(df$OP_dailymore, useNA = "ifany")


###METHADONE###
table(df$MD30, useNA = "ifany") #used methadone yes/no
table(df$MDNINJ, useNA = "ifany") #any non-injection methadone
table(df$MDX, useNA = "ifany") #non-injection methadone (# times)
table(df$MDINJ, useNA = "ifany") #any injection methadone 
table(df$MDIX, useNA = "ifany") #injection methadone (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$MDX <- as.integer(df$MDX) #change to numeric instead of factor
    df$MDIX <- as.integer(df$MDIX) #change to numeric instead of factor
    
    df$MD_dailymore <- ifelse(df$MDX>=30 | df$MDIX>=30, 1, 0)
    table(df$MD_dailymore, useNA = "ifany")

    
###BUPRENORPHINE###
table(df$BUP30, useNA = "ifany") #used bup yes/no
table(df$BUPNINJ, useNA = "ifany") #any non-injection bup 
table(df$BUPX, useNA = "ifany") #non-injection bup (# times)
table(df$BUPINJ, useNA = "ifany") #any injection bup 
table(df$BUPIX, useNA = "ifany") #injection bup (# times)

    #if we want to create a variable for at least daily speedball use (>=30 times past 30 days)
    df$BUPX <- as.integer(df$BUPX) #change to numeric instead of factor
    df$BUPIX <- as.integer(df$BUPIX) #change to numeric instead of factor
    
    df$BUP_dailymore <- ifelse(df$BUPX>=30 | df$BUPIX>=30, 1, 0)
    table(df$BUP_dailymore, useNA = "ifany")
```


Create variable for any stimulant use in past 30 days.
```{r}
####ANY STIMULANT USE IN PAST 30 DAYS (injection and/or non-injection composite)###
#includes crack cocaine, powder cocaine, meth, nonmedical rx stimulants, speedball, and goofball
df$ANYSTIMUSE <- ifelse(df$CRKNINJ=="Yes" | df$CCNINJ=="Yes" | df$MTNINJ=="Yes" | df$STNINJ=="Yes" |df$SBNINJ =="Yes" |df$GBNINJ=="Yes" | df$CRKINJ=="Yes" | df$CCINJ=="Yes" | df$MTINJ=="Yes" | df$STINJ=="Yes" |df$SBINJ=="Yes"| df$GBINJ=="Yes", 1, ifelse(df$CRKNINJ=="No" &df$CCNINJ=="No" & df$MTNINJ=="No" & df$STNINJ=="No" & df$SBNINJ=="No" & df$GBNINJ=="No" & df$CRKINJ=="No" &  df$CCINJ=="No" & df$MTINJ=="No" & df$STINJ=="No" & df$SBINJ=="No" & df$GBINJ=="No", 0, NA))

df$ANYSTIMUSE <- as.factor(df$ANYSTIMUSE)
table(df$ANYSTIMUSE, useNA = "ifany")
```

Create variable for any opioid use in past 30 days.
```{r}
###ANY OPIOID USE IN PAST 30 DAYS (injection and/or non-injection composite)###
#includes speedball, goofball, heroin, nonmedical rx opioids, methadone, bup
df$ANYOPIOIDUSE <- ifelse(df$SBINJ=="Yes" | df$GBINJ=="Yes" | df$MDINJ=="Yes" | df$BUPINJ=="Yes" | df$HRINJ=="Yes" | df$OPINJ=="Yes" |df$SBNINJ=="Yes" | df$GBNINJ=="Yes" |  df$MDNINJ=="Yes" |  df$BUPNINJ=="Yes" |  df$HRNINJ=="Yes" | df$OPNINJ=="Yes", 1, ifelse(df$SBINJ=="No" & df$GBINJ=="No"  & df$MDINJ=="No"  & df$BUPINJ=="No" & df$HRINJ=="No" & df$OPINJ=="No"  & df$SBNINJ=="No" & df$GBNINJ=="No"  & df$MDNINJ=="No"  & df$BUPNINJ=="No" & df$HRNINJ=="No" & df$OPNINJ=="No", 0, NA))

df$ANYOPIOIDUSE <- as.factor(df$ANYOPIOIDUSE)
table(df$ANYOPIOIDUSE, useNA = "ifany")
```

Create variable for any daily or more stimulant use in past 30 days.
```{r}
#ANY DAILY OR MORE STIMULANT USE (>=30 times in past 30 days)
df$STIM_dailymore <- ifelse(df$SB_dailymore==1 | df$GB_dailymore==1 | df$CRCK_dailymore==1 | df$PCOKE_dailymore==1 | df$METH_dailymore==1 | df$ST_dailymore==1, 1, 0)

table(df$STIM_dailymore, useNA = "ifany")
```

Create variable for any daily or more opioid use in past 30 days.
```{r}
#ANY DAILY OR MORE OPIOID USE (>=30 times in past 30 days)
df$OPIOID_dailymore <- ifelse(df$SB_dailymore==1 | df$GB_dailymore==1 | df$HER_dailymore==1 | df$OP_dailymore==1 | df$MD_dailymore==1 | df$BUP_dailymore==1, 1, 0)

table(df$OPIOID_dailymore, useNA = "ifany")
```

