---
title: "SER_oral_baseline"
author: "Elizabeth Kinnard"
date: "6/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval=F, warning = FALSE, message = FALSE, global.par=TRUE)
knitr::opts_knit$set(root.dir="~/Documents/GitHub/lizzyandshelley/Dataset")
```

Open packages.
```{r}
library(here)
library(foreign)
library(labelled)
library(dplyr)
library(reshape2)
library(lme4)
library(MASS)
library(mice)
library(biostat3)
library(lmtest)
library(gee)
library(ciTools)
library(forcats)
library(tableone)
```

Open working directory and load/label the data.
```{r}
#set working directory
setwd("~/Documents/GitHub/lizzyandshelley/Dataset") #LIZZY

#load the data
df <- read.spss("CTCAllWaves.SAV", use.value.labels = TRUE, to.data.frame=TRUE)

#get variable labels
attr(df, "variable.labels")
df.labels <- as.data.frame(attr(df, "variable.labels"))
QDSlabels <- data.frame(rownames(df.labels))
df.labels <- cbind(QDSlabels, df.labels)
names(df.labels) <- c("VAR_ID", "description")
rm(QDSlabels)

#USE THIS CODE TO LOOK UP A VARIABLE DESCRIPTION
#df.labels$description[df.labels$VAR_ID=="GENDER"]
```

Subsetting the analytic sample to people with greater than 0 injections and greater than never injecting with others. 
```{r}
#df.labels$description[df.labels$VAR_ID=="SUMINJ"]
#df.labels$description[df.labels$VAR_ID=="IJWOTH"]

#total number of injections last 30 days
df$SUMINJ <- as.numeric(as.character(df$SUMINJ))
table(df$SUMINJ, useNA = "ifany")

#frequency of injecting with others in last 6 months
table(df$IJWOTH, useNA = "ifany")

#convert to 3 level variable (0 = never, 1 = occasionally/sometimes, 2 = usually/always)
df$IJWOTH_3cat <- as.factor(ifelse(df$IJWOTH=="Never (0% of the time)", 0, ifelse(df$IJWOTH=="Occasionally (less than 25% of the time)" | df$IJWOTH=="Sometimes (25% to 74% of the time)", 1, ifelse(df$IJWOTH=="Usually (75% to 99% of the time)" | df$IJWOTH=="Always (100% of the time)", 2, NA))))

class(df$IJWOTH_3cat)
df$IJWOTH_3cat <- as.numeric(as.character(df$IJWOTH_3cat))
table(df$IJWOTH_3cat, useNA = "ifany")

#Subsetting the analytic sample to people with greater than 0 injections and greater than never injecting with others. 
df <- subset(df, SUMINJ>0) #sample from 984 to 979
df <- subset(df, IJWOTH_3cat>0) #sample from 979 to 879
```

Create public injection variable (exposure).
```{r}
#public injection
table(df$PUBINJ, useNA = "ifany") #Frequency of public injection (ordinal)

#convert to 3 level variable (0 = never, 1 = occasionally/sometimes, 2 = usually/always)
df$PUBINJ_3cat <- as.factor(ifelse(df$PUBINJ=="Never (0% of the time)", 0, ifelse(df$PUBINJ=="Occasionally (less than 25% of the time)" | df$PUBINJ=="Sometimes (25% to 74% of the time)", 1, ifelse(df$PUBINJ=="Usually (75% to 99% of the time)" | df$PUBINJ=="Always (100% of the time)", 2, NA))))

table(df$PUBINJ_3cat, useNA = "ifany") 
class(df$PUBINJ_3cat)
```

Create syringe sharing variable (outcome).
```{r}
table(df$RECSHR, useNA = "ifany") #Times used dirty needles last 6 months
df$RECSHR <- as.numeric(as.character(df$RECSHR))
```

Create covariates and substance use indicators.
```{r}
#age at baseline (years)
df$AGE <- as.numeric(as.character(df$AGE))
table(df$AGE, useNA = "ifany")

#gender (combine other and transgender into other so 3 cat variable)
table(df$GENDER, useNA = "ifany")
df$GENDER <- df$GENDER %>% fct_collapse("Other" = c("Transgender","Other"))

#homelessness
table(df$HOMELESS) #do you consider yourself homeless 
df$MOSHOME <- as.numeric(as.character(df$MOSHOME)) #number of months homeless
df$YRSHOME <- df$MOSHOME/12

table(df$HOMELESS, useNA = "ifany")
table(df$MOSHOME, useNA = "ifany")
table(df$YRSHOME, useNA = "ifany")

#crack injection
table(df$CRKINJ, useNA = "ifany")

#powder cocaine injection
table(df$CCINJ, useNA = "ifany")

#meth injection
table(df$MTINJ, useNA = "ifany")

#prescription stimulant injection
table(df$STINJ, useNA = "ifany")

#heroin injection
table(df$HRINJ, useNA = "ifany")

#NMPO injection
table(df$OPINJ, useNA = "ifany")

#alcohol
df$ALCDAY <- as.numeric(as.character(df$ALCDAY)) #Days of alcohol use in last 30 days
df$ALCTX <- as.numeric(as.character(df$ALCTX)) #Times you drank on a typical day
df$binge <- ifelse(df$ALCDAY==0, 0, ifelse(df$ALCTX<4, 0, ifelse(df$ALCTX>3, 1, NA))) #binge alcohol on a typical day (>3 drinks)

table(df$ALCDAY, useNA = "ifany")
table(df$ALCTX, useNA = "ifany")
table(df$binge, useNA = "ifany")

#any stimulant injection (composite)
df$ANYSTIM <- ifelse(df$CRKINJ=="Yes" | df$CCINJ=="Yes" | df$MTINJ=="Yes" | df$STINJ=="Yes", 1, ifelse(df$CRKINJ=="No" & df$CCINJ=="No" & df$MTINJ=="No" & df$STINJ=="No", 0, NA))
table(df$ANYSTIM, useNA = "ifany")

#any opioid injection (composite)
df$ANYOPIATE <- ifelse(df$HRINJ=="Yes" | df$OPINJ=="Yes", 1, ifelse(df$HRINJ=="No" & df$OPINJ=="No", 0, NA))
table(df$ANYOPIATE, useNA = "ifany")
```

Poisson model with no offset (and all covariates).
```{r}
modelpoisson <- glm(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, family = "poisson", data = df, na.action = na.omit)
summary(modelpoisson)
```

Poisson model with offset included (and all covariates).
```{r}
modelpoisson_offset <- glm(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, family = "poisson", offset = log(df$SUMINJ), data = df, na.action = na.omit)
summary(modelpoisson_offset)
```

Negative binomial model with no offset (and all covariates).
```{r}
model.nb <- glm.nb(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, data = df, na.action = na.omit)
summary(model.nb)
```

Negative binomial model offset included (and all covariates).
```{r}
model.nb.offset <- glm.nb(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE + offset(log(df$SUMINJ)), data = df, na.action = na.omit)
summary(model.nb.offset)

#confidence intervals for  coefficients
(CI_model.nb.offset <- cbind(Estimate = coef(model.nb.offset), confint(model.nb.offset)))
exp(CI_model.nb.offset)
```

FINAL MODEL 1: Negative binomial model offset included (homeless, gender, age included).
```{r}
finalmodel.nb1 <- glm.nb(RECSHR ~ PUBINJ_3cat + HOMELESS + GENDER + AGE + offset(log(df$SUMINJ)), data = df, na.action = na.omit)
summary(finalmodel.nb1)

#confidence intervals for  coefficients
(CI.finalmodel.nb1 <- cbind(Estimate = coef(finalmodel.nb1), confint(finalmodel.nb1)))

exp(CI.finalmodel.nb1)
```

FINAL MODEL 2: Negative binomial model offset included (only homelessness included as covariate).
```{r}
finalmodel.nb2 <- glm.nb(RECSHR ~ PUBINJ_3cat + HOMELESS + offset(log(df$SUMINJ)), data = df, na.action = na.omit)
summary(finalmodel.nb2)

#confidence intervals for  coefficients
(CI.finalmodel.nb2 <- cbind(Estimate = coef(finalmodel.nb2), confint(finalmodel.nb2)))

exp(CI.finalmodel.nb2)
```


Compare models.
```{r}
#comparing negative binomials with offset, with and without the anystim and anyopiate variables
lrtest(model.nb.offset, model.nb.offset.reduced)
AIC(model.nb.offset, model.nb.offset.reduced)
```

Interpretation:
- more like risk ratio because denominator is underlying number of events
- relative number of needle sharing incidents per injection 
- if exp(beta) = 2, twice the number of needle sharing events per injection 



```{r}
# Identify the variables of interest
myVars <- c("PUBINJ_3cat", "AGE", "GENDER", "HOMELESS", "RECSHR") 

# Specify which of the variables are categorical
catVars <- c("PUBINJ_3cat","GENDER","HOMELESS") 

# Create an overall table 1
table1.overall <- CreateTableOne(vars= myVars, 
                                 data= df, 
                                 factorVars = catVars)

# Create a table 1 stratified by smoking
table1 <- CreateTableOne(vars=myVars, 
                                   data=df, 
                                   factorVars = catVars)

print(table1, showAllLevels = TRUE)
```

