---
title: "SER_oral_baseline"
author: "Elizabeth Kinnard"
date: "4/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval=F, warning = FALSE, message = FALSE, global.par=TRUE)
knitr::opts_knit$set(root.dir="~/Documents/GitHub/lizzyandshelley/Dataset")
```

Open packages.
```{r}
library(here)
library(foreign)
library(labelled)
library(dplyr)
library(reshape2)
library(lme4)
library(MASS)
library(mice)
library(biostat3)
library(lmtest)
library(gee)
```

Open working directory and load/label the data.
```{r}
#set working directory
setwd("~/Documents/GitHub/lizzyandshelley/Dataset") #LIZZY
#setwd("~/GitHub/lizzyandshelley/Dataset")  #SHELLEY

#load the data
df <- read.spss("Ricky.SAV", use.value.labels = TRUE, to.data.frame=TRUE)

#get variable labels
attr(df, "variable.labels")
df.labels <- as.data.frame(attr(df, "variable.labels"))
QDSlabels <- data.frame(rownames(df.labels))
df.labels <- cbind(QDSlabels, df.labels)
names(df.labels) <- c("VAR_ID", "description")
rm(QDSlabels)

#USE THIS CODE TO LOOK UP A VARIABLE DESCRIPTION
#df.labels$description[df.labels$VAR_ID=="GENDER"]

#subsetting the data to only include people who injected in past 30 days (only removes 5 ppl)
df <- subset(df, SUMINJ>0)
```

Create public injection variable (exposure).
```{r}
#public injection
table(df$PUBINJ, useNA = "ifany") #Frequency of public injection (ordinal)

#convert to 3 level variable (0 = never, 1 = occasionally/sometimes, 2 = usually/always)
df$PUBINJ_3cat <- as.factor(ifelse(df$PUBINJ=="Never (0% of the time)", 0, ifelse(df$PUBINJ=="Occasionally (less than 25% of the time)" | df$PUBINJ=="Sometimes (25% to 74% of the time)", 1, ifelse(df$PUBINJ=="Usually (75% to 99% of the time)" | df$PUBINJ=="Always (100% of the time)", 2, NA))))

table(df$PUBINJ_3cat, useNA = "ifany") 
```

Create syringe sharing variable (outcome).
```{r}
table(df$RECSHR, useNA = "ifany") #Times used dirty needles last 6 months
df$RECSHR <- as.numeric(df$RECSHR)
```


Create covariates and substance use indicators.
```{r}
#age at baseline (years)
df$AGE <- as.numeric(df$AGE)
table(df$AGE, useNA = "ifany")

########IF HAVE TO CATEGORIZE AGE BECAUSE NB MODEL WOULDN'T CONVERGE###########
#df$AGE_cat <- as.factor(ifelse(df$AGE<25, "y", ifelse(df$AGE>24 & df$AGE<41, "m", ifelse(df$AGE>40, "e", NA))))

#gender
table(df$GENDER, useNA = "ifany")

#homelessness
df$HOMELESS <- as.numeric(df$HOMELESS) #do you consider yourself homeless
df$MOSHOME <- as.numeric(df$MOSHOME) #number of months homeless
df$YRSHOME <- df$MOSHOME/12

table(df$HOMELESS, useNA = "ifany")
table(df$MOSHOME, useNA = "ifany")
table(df$YRSHOME, useNA = "ifany")

#total number of injections last 30 days (offset term? should I keep this as 30 or multiply by 6?)
df$SUMINJ <- as.numeric(paste(df$SUMINJ))
table(df$SUMINJ, useNA = "ifany")

#frequency of injecting with others in last 6 months
table(df$IJWOTH, useNA = "ifany")

#crack injection
table(df$CRKINJ, useNA = "ifany")

#powder cocaine injection
table(df$CCINJ, useNA = "ifany")

#meth injection
table(df$MTINJ, useNA = "ifany")

#prescription stimulant injection
table(df$STINJ, useNA = "ifany")

#heroin injection
table(df$HRINJ, useNA = "ifany")

#NMPO injection
table(df$OPINJ, useNA = "ifany")

#alcohol
df$ALCDAY <- as.numeric(df$ALCDAY) #Days of alcohol use in last 30 days
df$ALCTX <- as.numeric(df$ALCTX) #Times you drank on a typical day
df$binge <- ifelse(df$ALCDAY==0, 0, ifelse(df$ALCTX<4, 0, ifelse(df$ALCTX>3, 1, NA))) #binge alcohol on a typical day (>3 drinks)

table(df$ALCDAY, useNA = "ifany")
table(df$ALCTX, useNA = "ifany")
table(df$binge, useNA = "ifany")

#any stimulant injection (composite)
df$ANYSTIM <- ifelse(df$CRKINJ=="Yes" | df$CCINJ=="Yes" | df$MTINJ=="Yes" | df$STINJ=="Yes", 1, ifelse(df$CRKINJ=="No" & df$CCINJ=="No" & df$MTINJ=="No" & df$STINJ=="No", 0, NA))
table(df$ANYSTIM, useNA = "ifany")

#any opioid injection (composite)
df$ANYOPIATE <- ifelse(df$HRINJ=="Yes" | df$OPINJ=="Yes", 1, ifelse(df$HRINJ=="No" & df$OPINJ=="No", 0, NA))
table(df$ANYOPIATE, useNA = "ifany")
```

Poisson model with no offset.
```{r}
modelpoisson <- glm(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, family = "poisson", data = df, na.action = na.omit)
summary(modelpoisson)
```

Poisson model with offset included.
```{r}
modelpoisson_offset <- glm(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, family = "poisson", offset = df$SUMINJ, data = df, na.action = na.omit)
summary(modelpoisson_offset)

#(WHY IS THERE AN ERROR HERE WHEN I TRY TO USE OFFSET: Error: no valid set of coefficients has been found: please supply starting values)
```

Negative binomial model with no offset.
```{r}
model.nb <- glm.nb(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, data = df, na.action = na.omit)
summary(model.nb)
```

Negative binomial model offset included.
```{r}
model.nb.offset <- glm.nb(RECSHR ~ PUBINJ_3cat*HOMELESS + GENDER + AGE + ANYSTIM + ANYOPIATE, data = df, offset= df$SUMINJ, na.action = na.omit)
summary(model.nb.offset)

#(WHY IS THERE AN ERROR HERE WHEN I TRY TO USE OFFSET:: Error in glm.control(...) : unused argument (offset = df$SUMINJ))
```

Compare poisson and negative binomial models.
```{r}
lrtest(modelpoisson, model.nb)
AIC(modelpoisson, model.nb)
```

Questions for Patrick:
1) Exposure as 3 level variable or keep it in its 5 level ordinal original form?

2) What to do with binge? (Not currently in model due to high degree of missingness)

3) What to do with injecting with others? (Conceptually does it make sense to include this and how to code, in its original form or same as exposure)

4) Zero inflated model? Because most people have zero on the outcome, but don't think the process was different for the zeros. "Furthermore, theory suggests that the excess zeros are generated by a separate process from the count values and that the excess zeros can be modeled independently."

5) What's going on with the offset conceptually and analytically?
